---
title: "Assignment 3"
author: "mutharasi"
date: "2024-08-29"
output: html_document
---

Assignment III  (Ex.No 11)
1. In this problem, you will develop a model to predict whether a given car gets high or low gas mileage based on the Auto data set.

(a) Create a binary variable, mpg01, that contains a 1 if mpg contains a value above its median, and a 0 if mpg contains a value below its median. You can compute the median using the median() function. Note you may find it helpful to use the data.frame() function to create a single data set containing both mpg01 and the other Auto variables.


```{r}
library(ISLR)

```

```{r}
summary(Auto)
```
```{r}
attach(Auto)
```
```{r}
mpg01 <- ifelse( mpg > median(mpg), 1, 0)
Auto <- data.frame(Auto, mpg01)

```

(b) Explore the data graphically in order to investigate the association between mpg01 and the other features. Which of the other features seem most likely to be useful in predicting mpg01? Scatterplots and boxplots may be useful tools to answer this question. Describe your findings. 

```{r}
cor(Auto[,-9])
```
cylinders, weight, displacement, horsepower are strong negatively correlated.

focusing on Weight, Displacement, and Horsepower will likely be most effective for predicting mpg01, as these features show the strongest associations with gas mileage.




```{r}
pairs(Auto)
```




(c)	 Split the data into a training set and a test set.

```{r}
train <-  (year %% 2 == 0) 
test <-  !train
Auto.train <-  Auto[train,]
Auto.test <-  Auto[test,]
mpg01.test <-  mpg01[test]
```


(d)	Perform LDA on the training data in order to predict mpg01 using the variables that seemed most associated with mpg01 in (b). What is the test error of the model obtained? 

```{r}
library(MASS)
lda.fit <-  lda(mpg01 ~ cylinders + weight + displacement + horsepower,
              data = Auto, subset = train)
lda.pred <-  predict(lda.fit, Auto.test)
mean(lda.pred$class != mpg01.test)
```
The test error of LDA model is 12.6%

(e)	Perform QDA on the training data in order to predict mpg01 using the variables that seemed most associated with mpg01 in (b). What is the test error of the model obtained? 


```{r}

qda.fit <- qda(mpg01 ~ cylinders + weight + displacement + horsepower, data = Auto, subset = train)
qda.pred <- predict(qda.fit, Auto.test)
mean(qda.pred$class != mpg01.test)

```
The test error of QDA model is 13.1%


(f)	Perform logistic regression on the training data in order to predict mpg01 using the variables that seemed most associated with mpg01 in (b). What is the test error of the model obtained?



```{r}
glm.fit <-  glm(mpg01 ~ cylinders + weight + displacement + horsepower,data = Auto,family = binomial,
              subset = train)
glm.probs <-  predict(glm.fit, Auto.test, type = "response")
glm.pred <-  rep(0, length(glm.probs))
glm.pred[glm.probs > 0.5] <- 1
mean(glm.pred != mpg01.test)
```
The test error of logistic regression model is 12.08%


(g)	Perform KNN on the training data, with several values of K, in order to predict mpg01. Use only the variables that seemed most associated with mpg01 in (b). What test errors do you obtain? Which value of K seems to perform the best on this data set?


```{r}
library(class)
train.X <-  cbind(cylinders, weight, displacement, horsepower)[train,]
test.X <-  cbind(cylinders, weight, displacement, horsepower)[test,]
train.mpg01 <-  mpg01[train]
set.seed(1)

#(k=1)
knn.pred <-  knn(train.X, test.X, train.mpg01, k = 1)
mean(knn.pred != mpg01.test)
```
The test error of KNN model when k=1 is 15.3%

```{r}
#(k=10)
knn.pred <-  knn(train.X, test.X, train.mpg01, k = 10)
mean(knn.pred != mpg01.test)
```
The test error of KNN model when k=10 is 16.4%

```{r}
#(k=100)
knn.pred <-  knn(train.X, test.X, train.mpg01, k = 100)
mean(knn.pred != mpg01.test)
```
The test error of KNN model when k=100 is 14.2%

k=100 is best because its error rate is lower then other.